---
title: "HW_4"
author: "Eugenia Lou"
date: "October 20, 2019"
output: html_document
---

HW_4a
```{r}
#setwd("c:/data/BUAN6357/HW_4");source("prep.txt", echo=T)
```

```{r}
library(tidyverse)
library(data.table)

# Delivery 1. seed: (vector) random number generator seed
seed <- 436604030
set.seed(seed)

aq  <- as.data.table(airquality)
```

```{r}
colnames(aq)
```

```{r}
head(aq)
```

```{r}
colSums(is.na(aq))
```

```{r}
raw <- na.omit(aq)                    # Remove missing value
raw <- subset(raw, select = -c(Day) ) # Delete column by column name
                                      # drop variable "Day"
raw$Month <- factor(raw$Month)        # Convert variable "Month" to factor

# Delivery 2. raw: (data.table) see Notes
raw
```

```{r}
# Delivery 3. base.model: (lm) non-CV OLS model using raw (all observations)
base.model <- lm(Ozone ~ ., data=raw)
summary(base.model)
```

```{r}
n <- nrow(raw)
m <- ceiling(0.1*nrow(raw))

# Delivery 4. tst: (vector) 10% validation set index values
tst   <- sample(n, m)
train <- raw[-tst, ]
test  <- raw[tst, ]
```

```{r}
# Delivery 5. cv.model: (lm) OLS model using simple CV
cv.model <- lm(Ozone ~ ., data=train)
summary(cv.model)
```

```{r}
# Delivery 6. cv.fitted: (vector) tst fitted values from cv.model; see Notes
cv.fitted <- predict.lm(cv.model,raw)[tst]
cv.fitted
```

```{r}
cv.resid <- NULL
num <- length(cv.fitted)

cv.resid$loc <- tst
cv.resid$diff <- test$Ozone-cv.fitted # residuals = actual - predicted
cv.resid <- as.data.table(cv.resid)

# Delivery 7. cv.resid (data.table) tst residuals from simple CV; see Notes
cv.resid
```

```{r}
# LOOCV (Leave one out cross validation)
loocv <- function(train.set, model){
  
  set.seed(seed)
  
  n  <- nrow(train.set)
  t  <- 1:n
  idx  <- sample(t, length(t))

  folds <- cut(seq(1, n), breaks=n ,labels=FALSE)
  
  loocv.predicted <- c()
  for(i in idx){
    #Segement the data by fold using the which() function 
    testIndexes <- which(folds==i, arr.ind=TRUE)
    testData    <- train.set[testIndexes, ]
    trainData   <- train.set[-testIndexes, ]
    
    lm.model  <- lm(model, trainData)
    predicted <- predict(lm.model, testData)
    loocv.predicted[i] <- predicted
    }
  
  return(loocv.predicted)
}

jk.resid <- NULL
n        <- nrow(raw)

loocv.predicted  <- loocv(raw, Ozone ~ .)

jk.resid$loc  <- 1:n
jk.resid$diff <- raw$Ozone-loocv.predicted # residuals = actual - predicted
jk.resid      <- as.data.table(jk.resid)

# Delivery 8. jk.resid: (data.table) validation residuals from LOOCV; see Notes
jk.resid
```

```{r}
# K-fold cross validation
n  <- nrow(raw)
k  <- 10
t <- rep( 1:k, each=ceiling(n/k) )
t <- t[1:n]

set.seed(seed)

idx  <- sample(n)
t2 <- data.frame(t,idx)

kf.resid <- data.table(k= 1:nrow(raw),loc=1:nrow(raw), diff=1:nrow(raw))
kf.resid$k <-t

tmp <- rep(0,ceiling(n/k))

kfold <- function(x) {
  tmp <- c()
  for (i in 1:x){
    kf.model<- lm(Ozone ~ ., data =raw[t2[t!=i,]$idx,] )
    kf.resid <- raw$Ozone[t2[t==i,]$idx] - predict(kf.model,raw)[t2[t==i,]$idx]
    tmp <- c(tmp,kf.resid)
  }
  return(tmp)
}

kf.resid$diff <-kfold(10)
kf.resid$k <- t2$t
kf.resid$loc <- t2$idx
kf.resid

```

```{r}
#source("validate.txt", echo=T)
```


HW_4b
Question-1
What is the average residual reported from the baseline model? (3 decimal places minimum)
```{r}
base.resid <- residuals(base.model)
summary(base.resid)
mean(base.resid)
```

Question-2
What is the parametric estimate of the lower bound for the 95% CI of the residuals reported from the baseline model? (3 decimal places minimum)
```{r}
# CI= [ xbar - z*std, xbar + z*std ]

z.value <- 1.96
std     <- sd(base.resid)

mean(base.resid)-z.value*std
```

Question-3
What is the non-parametric estimate of the upper bound for the 95% CI of the residuals reported from the baseline model? (3 decimal places minimum)
```{r}
# We do not assume distribution on non-parameter
quantile(base.resid, probs = 0.975)
```

Question-4
What is the average residual reported from the simple cross-validation model? (3 decimal places minimum)
```{r}
mean(cv.resid$diff)
```

Question-5
What is the parametric estimate of the lower bound for the 95% CI of the residuals reported from the simple cross-validation model? (3 decimal places minimum)
```{r}
z.value <- 1.96
std     <- sd(cv.resid$diff)

mean(cv.resid$diff)-z.value*std
```

Question-6
What is the non-parametric estimate of the upper bound for the 95% CI of the residuals reported from the simple cross-validation model? (3 decimal places minimum)
```{r}
quantile(cv.resid$diff, probs = 0.975)
```

Question-7
What is the average residual reported from the LOOCV cross-validation model? (3 decimal places minimum)
```{r}
mean(jk.resid$diff)
```

Question-8
What is the parametric estimate of the lower bound for the 95% CI of the residuals reported from the LOOCV cross-validation model? (3 decimal places minimum)
```{r}
z.value <- 1.96
std     <- sd(jk.resid$diff)

mean(jk.resid$diff)-z.value*std
```

Question-9
What is the non-parametric estimate of the upper bound for the 95% CI of the residuals reported from the LOOCV cross-validation model? (3 decimal places minimum)
```{r}
quantile(jk.resid$diff, probs = 0.975)
```

Question-10
What is the average residual reported from the K-fold (k=10) cross-validation model? (3 decimal places minimum)
```{r}
mean(kf.resid$diff)
```

Question-11
What is the parametric estimate of the lower bound for the 95% CI of the residuals reported from the K-fold (k=10) cross-validation model? (3 decimal places minimum)
```{r}
z.value <- 1.96
std     <- sd(kf.resid$diff)

mean(kf.resid$diff)-z.value*std
```

Question-12
What is the non-parametric estimate of the upper bound for the 95% CI of the residuals reported from the K-fold (k=10) cross-validation model? (3 decimal places minimum)
```{r}
quantile(kf.resid$diff, probs = 0.975)
```


