---
title: "HW_5"
author: "Eugenia Lou"
date: "October 21, 2019"
output: html_document
---

HW_5a
```{r}
#setwd("c:/data/BUAN6357/HW_5");source("prep.txt", echo=T)
```

```{r}
library(tidyverse)
library(broom)
library(data.table)

seed <- 436604030
set.seed(seed)

aq  <- as.data.table(airquality)
```

```{r}
raw <- na.omit(aq)                    # Remove missing value
raw <- subset(raw, select = -c(Day) ) # Delete column by column name
                                      # drop variable "Day"
raw$Month <- factor(raw$Month)        # Convert variable "Month" to factor
raw
```

```{r}
n <- nrow(raw)  # number of observations in raw
b <- 500        # number of bootstrap iterations
```

```{r}
# Classic bootstrap- for loop
set.seed(seed)

t     <- NULL
g     <- NULL
idx   <- NULL
cl.loop <- NULL

t <- as.vector(replicate(b, sample(n, replace=T), simplify=T)) 
g <- rep(1:b, each=n)

for (i in 1:b ){
  idx     <- t[g==i] # index values for sample
  cl.loop <- as.data.table(rbind(cl.loop, tidy(lm(Ozone~ ., data= raw[idx]))))
}

cl.loop
```

```{r}
# Classic bootstrap- for DT
set.seed(seed)

t   <- NULL
g   <- NULL
dt  <- NULL
cl.dt <- NULL
 
t <- as.vector(replicate(b, sample(n, replace=T), simplify=T))
g <- rep(1:b, each=n)

dt    <- data.table(t= t, g= g)
cl.dt <- dt[,tidy(lm(Ozone~ ., data= raw[t])), by= g]
cl.dt

#all.equal(cl.dt, cl.loop)
```

```{r}
# Balanced bootstrap- for loop
set.seed(seed)

t   <- NULL
g   <- NULL
idx <- NULL
bal.loop <- NULL

t <- sample(rep(1:n, b))
g <- rep(1:b, each=n)

for (j in 1:b ){
  idx <- t[g==j] # index values for sample
  bal.loop <- as.data.table(rbind(bal.loop, tidy(lm(Ozone~ ., data = raw[idx]))))
}

bal.loop

```

```{r}
# Balanced bootstrap- for DT
set.seed(seed)

t   <- NULL
g   <- NULL
dt  <- NULL
bal.dt <- NULL

t <- sample(rep(1:n, b))
g <- rep(1:b, each=n)

dt <- data.table(t= t, g= g)
bal.dt <- dt[,tidy(lm(Ozone~ ., data = raw[t])), by= g]
bal.dt

#all.equal(bal.dt, bal.loop)
```

```{r}
#source("validate.txt", echo=T)
```

HW_5b
Question-1
What is the average value for the Wind estimate as calculated using the Classic Bootstrap approach results? (3 digits minimum)
```{r}
wind.cl <- filter(cl.dt, term=="Wind")
mean(wind.cl$estimate)
```

Question-2
What is the non-parametric 95% CI lower bound for the Wind estimate calculated from the Classic Bootstrap approach results? (3 digits minimum)
```{r}
quantile(wind.cl$estimate, probs = 1-0.975)
```

Question-3
What is the parametric 95% CI upper bound for the Wind estimate calculated from the Classic Bootstrap approach results? (3 digits minimum)
```{r}
z.value <- 1.96
std     <- sd(wind.cl$estimate)

mean(wind.cl$estimate)+z.value*std
```

Question-4
What is the standard deviation value for the Temp estimate as calculated using the Classic Bootstrap approach results? (3 digits minimum)
```{r}
temp.cl <- filter(cl.dt, term=="Temp")
sd(temp.cl$estimate)
```

Question-5
What is the non-parametric 95% CI upper bound for the Temp estimate calculated from the Classic Bootstrap approach results? (3 digits minimum)
```{r}
quantile(temp.cl$estimate, probs = 0.975)
```

Question-6
What is the parametric 95% CI lower bound for the Temp estimate calculated from the Classic Bootstrap approach results? (3 digits minimum)
```{r}
z.value <- 1.96
std     <- sd(temp.cl$estimate)

mean(temp.cl$estimate)-z.value*std
```

Question-7
What is the proportion of p.values less than alpha=0.05 for the Wind estimate when using the Classic Bootstrap approach results? (0<=X<=1, 3 digits minimum)
```{r}
# Probability(p-value<=0.05)= ?
alpha <- 0.05
sum(wind.cl$p.value<=alpha)/length(wind.cl$p.value)
```

Question-8
What is the standard deviation value for the Wind estimate as calculated using the Balanced Bootstrap approach results? (3 digits minimum)
```{r}
wind.bal <- filter(bal.dt, term=="Wind")
sd(wind.bal$estimate)
```

Question-9
What is the non-parametric 95% CI upper bound for the Wind estimate calculated from the Balanced Bootstrap approach results? (3 digits minimum)
```{r}
quantile(wind.bal$estimate, probs = 0.975)
```

Question-10
What is the parametric 95% CI lower bound for the Wind estimate calculated from the Balanced Bootstrap approach results? (3 digits minimum)
```{r}
z.value <- 1.96
std     <- sd(wind.bal$estimate)

mean(wind.bal$estimate)-z.value*std
```

Question-11
What is the average value for the Temp estimate as calculated using the Balanced Bootstrap approach results? (3 digits minimum)
```{r}
temp.bal <- filter(bal.dt, term=="Temp")
mean(temp.bal$estimate)
```

Question-12
What is the non-parametric 95% CI lower bound for the Temp estimate calculated from the Balanced Bootstrap approach results? (3 digits minimum)
```{r}
quantile(temp.bal$estimate, probs = 1-0.975)
```

Question-13
What is the parametric 95% CI upper bound for the Temp estimate calculated from the Balanced Bootstrap approach results? (3 digits minimum)
```{r}
z.value <- 1.96
std     <- sd(temp.bal$estimate)

mean(temp.bal$estimate)+z.value*std
```

Question-14
What is the proportion of p.values less than alpha=0.05 for the Temp estimate when using the Balanced Bootstrap approach results? (0<=X<=1, 3 digits minimum)
```{r}
# Probability(p-value<=0.05)= ?
alpha <- 0.05
sum(temp.bal$p.value<=alpha)/length(temp.bal$p.value)
```



